---
title: "Working with Bonds"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Bonds}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| echo: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Introduction

In Formal Concept Analysis (FCA), a **bond** between two formal contexts $K_1 = (G_1, M_1, I_1)$ and $K_2 = (G_2, M_2, I_2)$ is a relation $R \subseteq G_1 \times M_2$ such that:

- For every object $g \in G_1$, the set of attributes $\{m \in M_2 \mid (g, m) \in R\}$ is an intent of $K_2$.
- For every attribute $m \in M_2$, the set of objects $\{g \in G_1 \mid (g, m) \in R\}$ is an extent of $K_1$.

Bonds represent "compatible" relations between the objects of one context and the attributes of another. Mathematically, the set of all bonds between two contexts, ordered by inclusion, forms a complete lattice called the **Bond Lattice**.

In `fcaR`, bonds are treated as first-class citizens with a dedicated `BondLattice` class that extends the standard `ConceptLattice`.

```{r}
library(fcaR)
```

# Computing Bonds

The main function to compute bonds is `bonds()`. It takes two `FormalContext` objects as input.

To keep this example fast, we will generate two small random formal contexts ($4 \times 4$).

```{r}
set.seed(42)
# Context 1
mat1 <- matrix(sample(0:1, 16, replace = TRUE), nrow = 4, ncol = 4)
rownames(mat1) <- paste0("O", 1:4)
colnames(mat1) <- paste0("A", 1:4)
fc1 <- FormalContext$new(mat1)
print(fc1)

# Context 2
mat2 <- matrix(sample(0:1, 16, replace = TRUE), nrow = 4, ncol = 4)
rownames(mat2) <- paste0("P", 1:4)
colnames(mat2) <- paste0("B", 1:4)
fc2 <- FormalContext$new(mat2)
print(fc2)
```


To compute the bond lattice:

```{r}
bl <- bonds(fc1, fc2, method = "conexp")
bl
```


## Computation Methods

The `bonds()` function provides two optimized C++ methods:

1. `"conexp"` (Default): Uses an implication-based approach on a tensor product of the contexts. It is generally the fastest for dense or moderately sized contexts.
2. `"mcis"`: A backtracking algorithm that operates directly on the pre-computed concept sets of both contexts. It can be more efficient in specific structural configurations.

```{r}
# Using the backtracking method
bl_mcis <- bonds(fc1, fc2, method = "mcis")
bl_mcis$size()
```


# The BondLattice Object

The result of `bonds()` is an object of class `BondLattice`. Since this class inherits from `ConceptLattice`, you can use all standard lattice operations.

## Visualization

You can plot the Hasse diagram of the bond lattice:

```{r
bl$plot()
```

## Extracting Bonds

Each node in the bond lattice represents a specific bond (a relation). You can extract these relations as individual `FormalContext` objects:

```{r}
# Get all bonds as a list of FormalContexts
all_bonds <- bl$get_bonds()
length(all_bonds)

# Inspect the first non-trivial bond
# (Note: Bond 1 is usually the "Core" or minimal bond)
all_bonds[[1]]
```


## The Core Bond

The "Core" bond is the intersection of all bonds (the infimum of the lattice). It represents the most fundamental consensus between the two contexts.

```{r}
core <- bl$get_core()
core$plot()
```


# Verifying Bonds

If you have a relation (as a matrix or a `FormalContext`) and want to check if it satisfies the mathematical definition of a bond between two contexts:

```{r}
# Take an existing bond and check it
mat_bond <- methods::as(all_bonds[[1]]$incidence(), "matrix")
is_bond(fc1, fc2, mat_bond)

# Check an arbitrary (likely invalid) relation
random_rel <- matrix(0, nrow = nrow(mat_bond), ncol = ncol(mat_bond))
is_bond(fc1, fc2, random_rel)
```


# Similarity and Complexity Metrics

The `BondLattice` class provides a `similarity()` method to compute various metrics that describe the relationship between the two formal contexts.

The following metrics are available. Let $L_{12}$ be the bond lattice between $K_1$ and $K_2$, and $L_{11}$ and $L_{22}$ be the bond lattices of $K_1$ and $K_2$ with themselves, respectively. We denote the size (number of concepts) of a lattice $L$ as $|L|$.

- `log-bond`: Measures how much the two contexts share a common logical structure. It is calculated as the normalized log-ratio of bonds:
  $$ \text{Log-Bond} = \frac{\log(|L_{12}|)}{\sqrt{\log(|L_{11}|) \cdot \log(|L_{22}|)}} $$
- `complexity`: Ratio of irreducible bonds to total bonds. Lower values indicate more emergent structural properties. Let $JI(L_{12})$ be the set of join-irreducible elements of the bond lattice:
  $$ \text{Complexity} = \frac{|JI(L_{12})|}{|L_{12}|} $$
- `core-agreement`: Ratio of filled cells in the Core bond versus the Top (largest) bond. If $B_{core}$ is the Core bond and $B_{top}$ is the Top bond, and $|B|$ represents the number of elements in the relation (filled cells):
  $$ \text{Core-Agreement} = \frac{|B_{core}|}{|B_{top}|} $$
- `entropy`: Based on the log-size of the lattices. It measures interaction entropy as:
  $$ \text{Entropy} = \frac{\log(|L_{12}|)}{\log(|L_{11}|) + \log(|L_{22}|)} $$


The `similarity()` method returns a named vector of these metrics.

```{r}
# 1. Logical Affinity (Log-Bond)
# Measures how much the two contexts share a common logical structure.
# 1.0 means perfect affinity.
bl$similarity("log-bond")

# 2. Structural Complexity
# Ratio of irreducible bonds to total bonds.
# Lower values indicate more emergent structural properties.
bl$similarity("complexity")

# 3. Core Agreement
# Ratio of filled cells in the Core bond versus the Top (largest) bond.
bl$similarity("core-agreement")

# 4. Interaction Entropy
# Based on the log-size of the lattices.
bl$similarity("entropy")
```


# Order-Theoretic Properties

Bonds also allow exploring deep structural properties of the interaction between contexts using measures like **Width** and **Dimension**.

- **Dilworth's Width**: The size of the largest antichain in the bond lattice.
- **Order Dimension**: The minimum number of linear orders whose intersection is the bond lattice.

```{r}
# Dilworth's Width
bl$similarity("width")

# Order Dimension (estimated via heuristic)
bl$similarity("dimension")
```


These measures can also be expressed as indices normalized by the lattice size:

- **Width Index**:
  $$ \text{Width-Index} = \frac{\text{Width}(L_{12})}{|L_{12}|} $$
- **Dimension Index**:
  $$ \text{Dimension-Index} = \frac{\text{Dimension}(L_{12})}{\log_2(|L_{12}|)} $$

```{r}
bl$similarity("width-index")
bl$similarity("dimension-index")
```


# Summary

Bonds provide a powerful mathematical framework to analyze the alignment or interaction between different perspectives (contexts) on the same objects or attributes. With `fcaR`, you can efficiently calculate large bond lattices, visualize them, and extract meaningful metrics to quantify context similarity and structural emergence.
