---
title: "Simplification Logic for Mushroom Dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mushroom}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fcaR)
```

# Introduction

This is a simple example of some of the functionalities introduced in package `fcaR`:

- Import from/export to `arules` format.
- Removal of redundancies in implications.
- Computation of closures.

# Data

In this example, we'll use the well-known [_Mushroom_ dataset](http://archive.ics.uci.edu/ml/datasets/Mushroom), from the `arules` package. 

We'll use the _a priori_ algorithm to extract a large number of implications from the dataset.

```{r}
library(arules)

data("Mushroom", package = "arules")

mush <- apriori(Mushroom, parameter = list(conf = 1))
```

The number of implications extracted by the algorithm (with this configuration) is `length(mush) = ` `r length(mush)`.

# Preprocessing

Next step is to remove the redundant rules in the implications set. To this end, let us use the `is.redundant()` function in `arules`.

```{r}
system.time(
  idx_redundant <- is.redundant(mush)
)

mush_clean <- mush[!idx_redundant]
```
After this, the number of implications is `length(mush_clean) = ` `r length(mush_clean)`, that is, just a `r round(length(mush_clean) / length(mush) * 100, 3)` percent of the original ruleset.

# Importing in fcaR

To use the functionalities in this package, one must import all objects (the formal context and implications above) into our data model.

This is accomplished by just typing:
```{r}
fc <- formal_context$new(I = Mushroom)

fc$add_implications(mush_clean)
```

We can check some of the properties of the ruleset:

```{r}
# Cadinality
fc$implications$cardinality()

# Rule size
sizes <- fc$implications$size()

# Mean size for LHS and RHS
colMeans(sizes)
```

# Applying Rules

We can improve the redudancy removal performed by `arules`, using some simplification rules. Currently, the following rules are implemented:

- Composition.
- Generalization.
- Simplification.

We can apply them one by one, just to see their effect, or we could use them sequentially.

```{r}
# Use composition to reduce the number of implications
fc$implications$apply_rules(rules = c("composition"))

fc$implications$cardinality()
sizes <- fc$implications$size()
colMeans(sizes)

# Simplification
fc$implications$apply_rules(rules = c("simplification"))

fc$implications$cardinality()
sizes <- fc$implications$size()
colMeans(sizes)

# At this moment, we're at a fixed point, but we could apply
# some more rules if needed:
fc$implications$apply_rules(rules = c("composition",
                                      "generalization",
                                      "simplification"))
```

Also, we can compute the support of each implication:
```{r}
supp <- fc$get_implication_support()

head(supp)
```


# Computing Closure

Given a fuzzy set representd by a sparse matrix, we can compute its closure wrt the implications in the ruleset.

```{r}
# LHS of the fifth rule
A <- fc$implications$get_LHS_matrix()[, 5]
# Associated attributes
print_set(A, fc$attributes)

# Compute the closure
cl <- fc$implications$compute_closure(A)
# Associated attributes
print_set(cl, fc$attributes)
```


# Re-Exporting to `arules`

After this phase of redudancy removal, we can export the obtained ruleset to `arules format`.

```{r}
R <- fc$convert_implications_to_arules()

R

class(R)
```

