---
title: "Simplification Logic for Fuzzy and Crisp Datasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simplification Logic for Fuzzy and Crisp Datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fcaR)
```

# Introduction

This is a simple example of some of the functionalities introduced in package `fcaR`:

- Computing implications and concepts using Ganter's algorithm.
- Visualization of the concept lattice.
- Removal of redundancies in implications.
- Computation of closures.

# Data

The datasets in this vignette come from [this paper](https://www.sciencedirect.com/science/article/pii/S1877705812021418).

# Crisp Version

The crisp version of the data appears in Table 3 in the mentioned paper.

```{r}
objects <- paste0("O", 1:6)
n_objects <- length(objects)

attributes <- paste0("P", 1:6)
n_attributes <- length(attributes)

I <- matrix(data = c(0, 1, 1, 0, 0, 1,
                     1, 1, 1, 0, 0, 0,
                     1, 1, 0, 0, 1, 0,
                     1, 0, 0, 1, 1, 0,
                     1, 0, 0, 1, 0, 0,
                     0, 0, 1, 0, 0, 0),
            nrow = n_objects,
            byrow = FALSE)

colnames(I) <- attributes
rownames(I) <- objects
```

## Computing Implications and Concepts

Once we create the formal context object, with the previous data matrix I, we can compute all concepts and implications using Ganter's algorithm:
```{r}
fc <- FormalContext$new(I)

# Compute implications
fc$extract_implications_concepts(verbose = FALSE)

# Cardinality and mean size in the ruleset
fc$implications$cardinality()
sizes <- fc$implications$size()
colMeans(sizes)
```

The obtained implications are:
```{r}
fc$implications
```

## Visualization

We provide functions to plot both the concept lattice and the formal context:

```{r fig.width=7.5}
# Visualize the concept lattice
fc$plot_lattice()
# And the formal context
fc$plot_context()
```

## Redudancy Removal

Let us apply some simplifcation rules:
```{r}
fc$implications$apply_rules(rules = c("composition",
                                      "generalization",
                                      "simplification"),
                            parallelize = FALSE)

# Compute cardinality and size in the transformed ruleset:
fc$implications$cardinality()
sizes <- fc$implications$size()
colMeans(sizes)
```

The transformed ruleset:
```{r}
fc$implications
```

# Fuzzy version

The fuzzy version of the data appears in Table 2 in the mentioned paper.

```{r}
objects <- paste0("O", 1:6)
n_objects <- length(objects)

attributes <- paste0("P", 1:6)
n_attributes <- length(attributes)

I <- matrix(data = c(0, 1, 0.5, 0, 0, 0.5,
                     1, 1, 0.5, 0, 0, 0,
                     0.5, 1, 0, 0, 1, 0,
                     0.5, 0, 0, 1, 0.5, 0,
                     1, 0, 0, 0.5, 0, 0,
                     0, 0, 1, 0, 0, 0),
            nrow = n_objects,
            byrow = FALSE)

colnames(I) <- attributes
rownames(I) <- objects
```

## Computing Implications and Concepts

As before, we build the formal context object and compute all implications:

```{r}
fc <- FormalContext$new(I)

# Compute
fc$extract_implications_concepts(verbose = FALSE)

# Some properties of the ruleset
fc$implications$cardinality()
sizes <- fc$implications$size()
colMeans(sizes)
```

The extracted ruleset is:
```{r}
fc$implications
```

## Visualization

Let us plot the concept lattice and the formal context.
```{r fig.width=7.5}
# Visualize the concept lattice
fc$plot_lattice()
# And the formal context
fc$plot_context()
```

## Redudancy Removal

Let us apply some functions to remove redudancies in the set of implications:

```{r}
fc$implications$apply_rules(rules = c("composition"), 
                            reorder = FALSE,
                            parallelize = FALSE)
fc$implications$cardinality()
sizes <- fc$implications$size()
colMeans(sizes)

fc$implications$apply_rules(rules = c("simplification"), 
                            reorder = FALSE,
                            parallelize = FALSE)
fc$implications$cardinality()
sizes <- fc$implications$size()
colMeans(sizes)

fc$implications$apply_rules(rules = c("generalization"), 
                            reorder = FALSE,
                            parallelize = FALSE)

fc$implications$cardinality()
sizes <- fc$implications$size()
colMeans(sizes)
```

The reduced ruleset is:
```{r}
fc$implications
```

## Closures

Let us show how to compute the closure of a set S with respect to the implication set.
```{r}
S <- SparseSet$new(attributes = fc$attributes)
S$assign(attributes = c("P2", "P3"),
         values = c(0.5, 0.5))
S

fc$implications$compute_closure(S)
```

